Prognostic and biologic relevance of collagen biosynthesis pathway in prostate cancer

### 2023-07-27 report/markdown/manuscript_figures_tables.Rmd
Supplementary material backbone
First supplementary figures
Supplementary figures for the normal - tumor comparison, cluster definition, infiltration prediction, GSVA and differential gene expression

_(A) Mean normalized $log_2$-transformed expression levels of the collagen pathway genes in the collagen clusters and cohorts presented in a heat map._

@@

_(A) Counts of cancer-associated fibroblasts (CAF) and endothelial cells (EC) estimated by the MCP counter algorithm were found to differ significantly between the collagen clusters in all investigated collectives. Statistical significance was assessed by Kruskal-Wallis test with $\eta^2$ effect size statistic. P values were corrected for multiple comparison with the false discovery rate method. Counts of fibroblasts and endothelial cells in the collagen clusters in the TCGA cohort are shown in violin plots. Red diamonds with whiskers represented medians with interquartile ranges. Single tumor samples are visualized as points. Effect sizes and p values are displayed in the plot captions. Numbers of observations in the clusters are indicated in the X axes._

_(B) Gene set variation analysis (GSVA) with Reactome pathway gene signatures. Statistical significance for differences in signature single sample gene set enrichment analysis scores (ssGSEA) between the collagen clusters was determined by one-way ANOVA with $\eta^2$ effect size statistic. P values were corrected for multiple testing with the false discovery rate method. Mean ssGSEA scores in the clusters and cohorts for signatures significantly regulated in at least four cohorts are presented in a heat map. Full analysis results are presented in Supplementary Table S6._

_(C) Modulation of KEGG-listed signaling pathways in the collagen^high^ (hi) or collagen^intermediate^ (hi) as compared with collagen^low^ tumors was predicted by the SPIA algorithm based on differential gene expression estimates. Predicted magnitude of pathway modulation (tA parameter) for pathways significantly modulated in at least four cohorts was visualized in bubble plots (tA < 0: inhibition, blue; tA > 0, activation, red). Points are labeled with their tA values. Significant effects are highlighted in bold. Full analysis results are presented in Supplementary Table S8._

_(D) Modulation of Recon2 model metabolic reactions in the collagen^high^ (hi) or collagen^intermediate^ (hi) as compared with collagen^low^ tumors was predicted by the BiGGR and biggrExtra algorithms based on differential gene expression estimates. Enrichment of significantly activated and significantly inhibited reactions in the collagen^high^ or collagen^intermediate^ clusters within the Recon model metabolic subsystems was investigated by Fisher's exact test. Enrichment p values were corrected for multiple testing with the false discovery rate method. Enrichment odds ratios for activated (red) and inhibited reaction (blue) for significantly enriched metabolic subsystems shared by at least four cohorts are presented in bubble plots. Points are labeled with their odds ration values. Full analysis results are presented in Supplementary Table S10._

>>>>

_(A) Mean normalized $log_2$-transformed expression levels of the collagen pathway genes in the collagen clusters and cohorts presented in a heat map. ECM: extracellular matrix._

@@

_(A) Counts of cancer-associated fibroblasts (CAF) and endothelial cells (EC) estimated by the MCP counter algorithm were found to differ significantly between the collagen clusters in all investigated collectives. Statistical significance was assessed by Kruskal-Wallis test with $\eta^2$ effect size statistic. P values were corrected for multiple comparison with the false discovery rate method. Counts of fibroblasts and endothelial cells in the collagen clusters in the TCGA cohort are shown in violin plots. Red diamonds with whiskers represented medians with interquartile ranges. Single tumor samples are visualized as points. Effect sizes and p values are displayed in the plot captions. Numbers of observations in the clusters are indicated in the X axes. Full analysis results are presented in Supplementary Table S4._

_(B) Gene set variation analysis (GSVA) with Reactome pathway gene signatures. Statistical significance for differences in signature single sample gene set enrichment analysis scores (ssGSEA) between the collagen clusters was determined by one-way ANOVA with $\eta^2$ effect size statistic. P values were corrected for multiple testing with the false discovery rate method. Mean ssGSEA scores in the clusters and cohorts for signatures significantly regulated in at least four cohorts are presented in a heat map. Full analysis results are presented in Supplementary Table S6. ECM: extracellular matrix, GF: growth factor._

_(C) Modulation of KEGG-listed signaling pathways in the collagen^high^ (hi) or collagen^intermediate^ (hi) as compared with collagen^low^ tumors was predicted by the SPIA algorithm based on differential gene expression estimates. Predicted magnitude of pathway modulation (tA parameter) for pathways significantly modulated in at least four cohorts was visualized in bubble plots (tA < 0: inhibition, blue; tA > 0, activation, red). Points are labeled with their tA values. Significant effects are highlighted in bold. Full analysis results are presented in Supplementary Table S8. ECM: extracellular matrix, ACT: actin, SCLC: small cell lung carcinoma, S. aureus: Staphylococcus aureus, SLE: systemic lupus erythematosus._

_(D) Modulation of Recon2 model metabolic reactions in the collagen^high^ (hi) or collagen^intermediate^ (hi) as compared with collagen^low^ tumors was predicted by the BiGGR and biggrExtra algorithms based on differential gene expression estimates. Enrichment of significantly activated and significantly inhibited reactions in the collagen^high^ or collagen^intermediate^ clusters within the Recon model metabolic subsystems was investigated by Fisher's exact test. Enrichment p values were corrected for multiple testing with the false discovery rate method. Enrichment odds ratios for activated (red) and inhibited reaction (blue) for significantly enriched metabolic subsystems shared by at least four cohorts are presented in bubble plots. Points are labeled with their odds ration values. Full analysis results are presented in Supplementary Table S10. Val: valine, Leu: leucine, Ile: isoleucine, FAOX: fatty acid oxidation._

### 2023-07-27 report/markdown/manuscript_supplement.Rmd
First supplementary figures
Supplementary figures for the normal - tumor comparison, cluster definition, infiltration prediction, GSVA and differential gene expression

flextable::flextable(suppl_paper_tbl$clinic) %>% 
  set_widths(c(2.2, 3.5, rep(4.5, 3), 2.8, 2.8)) %>% 
  merge_v(1:2) %>% 
  footnote(1, 2, 
           value = as_paragraph('PSA: prostate-specific antigen.'), 
           part = 'header', 
           ref_symbols = 'a') %>% 
  footnote(1, 6:7, 
           value = as_paragraph("Categorical variables: \u03C7\u00B2 test with Cramer's V effec size statistic. Numeric variables: Kruskal-Wallis test with \u03B7\u00B2 effect size statistic. P values corrected for multiple testing with the false discovery rate method."), 
           part = 'header', 
           ref_symbols = 'b') %>% 
  theme_vanilla

>>>>

flextable::flextable(suppl_paper_tbl$clinic) %>% 
  set_widths(c(2.2, 3, rep(3.5, 3), 2.8, 2.8)) %>% 
  merge_v(1:2) %>% 
  footnote(1, 2, 
           value = as_paragraph('PSA: prostate-specific antigen.'), 
           part = 'header', 
           ref_symbols = 'a') %>% 
  footnote(1, 6:7, 
           value = as_paragraph("Categorical variables: \u03C7\u00B2 test with Cramer's V effec size statistic. Numeric variables: Kruskal-Wallis test with \u03B7\u00B2 effect size statistic. P values corrected for multiple testing with the false discovery rate method."), 
           part = 'header', 
           ref_symbols = 'b') %>% 
  theme_vanilla
